---
title: "Capstone Project"
author: "WaleAOT"
date: "4/8/2022"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---


```{r setup, include=FALSE}
#bring in the artilleries
library(tidyverse)
library(tidyr)
library(readr)
library(dplyr)
library(flexdashboard)
library(stringr)
library(plotly)
```


```{r}
#bring in the datasets
orders <- read_csv("orders.csv")
people <- read_csv("people.csv")
returns <- read_csv("returns.csv")

```



page 1
==============================================================================

Row
------------------------------------------------------------------------------



### Table of Profits per country for 2014

```{r}
orders <- orders %>% mutate(Year = str_sub(Ship_Date,-4,-1))


highest_country <- orders %>%
                filter(Year == "2014") %>%
            #extract profits by country
                   group_by(Country,Year)%>%
   summarise(Total_profit = sum(round(Profit,digits = 4)))


highest_country <- highest_country %>% 
                  as.data.frame() %>% 
#rearrange profit from highest to lowest(since we are concerned with highest                            values)
                    arrange(desc(Total_profit))

highest_country

```
Table of profit per country


              
```{r}
#focusing on the top 20 countries for convenience...
  highest_country_top_20 <- highest_country[c(1:30),]

#Let ggplot order the plot by total_profit
  highest_country_top_20$Country <- factor(highest_country_top_20$Country, 
                                  levels =   highest_country_top_20$Country[order(highest_country_top_20$Total_profit)])
  
  
```



page 2
=======================================================================

Row
-----------------------------------------------------------------------

### Plots of countries by Total profit in 2014. 

```{r echo=FALSE, message=FALSE, paged.print=FALSE}
#Visualization proper

q1 <- ggplot(data = highest_country_top_20, aes(x= Total_profit, y=Country, fill= Country)) +   geom_bar(stat="identity") + ggtitle(label = "Plots of countries by Total profit")+ geom_text(aes(label = round(Total_profit)),hjust = -0.1)
ggplotly(q1)

```




From the plot, United States of America, China and India are the top 3 countries with the highest profit in the year 2014.


Question 1b
 For each of these three countries, find the three products with the highest total profit.
Specifically, what are the products’ names and the total profit for each product?



```{r}
# Getting the products for the top 3 countries
#capture product name and profit for the 3 countries

Top_Country <- orders %>%
                select(Country, Profit,Product_Name, Year) %>%
                filter(Year == "2014" & Country %in%  c("United States",  "India" , "China"))
head(Top_Country)

```

page 3
==============================================================================

Row
------------------------------------------------------------------------------

### Top 3 products in the USA with highest Total Profit

```{r}
#Get the data per country
#for USA
Top_Country_US <- Top_Country %>%
                filter( Country == "United States") %>%
                group_by(Product_Name) %>%
                summarise(Total_profit = sum(Profit))

#Convert the tibble to a dataframe, so we can rank Total_profit by desc.

Top_Country_US_df <- Top_Country_US %>%
                      as.data.frame() %>%
                      arrange(desc(Total_profit))

head(Top_Country_US_df,3)

```

Row
-----------------------------------------------------------------------------

### Top 3 products in India with highest Total Profit


```{r}
Top_Country_India <- Top_Country %>%
                filter( Country == "India") %>%
                group_by(Product_Name) %>%
                summarise(Total_profit= sum(Profit))

#Convert the tibble to a dataframe, so we can rank.

Top_Country_India_df <- Top_Country_India %>%
                      as.data.frame() %>%
                      arrange(desc(Total_profit))

head(Top_Country_India_df,3)
```

Row
------------------------------------------------------------------------------

### Top 3 products in China with highest Total Profit

```{r}

#For China
Top_Country_China <- Top_Country %>%
                filter( Country == "China") %>%
                group_by(Product_Name) %>%
                summarise(Total_profit= sum(Profit))

#Convert the tibble to a dataframe, so we can rank.

Top_Country_china_df <- Top_Country_China %>%
                      as.data.frame() %>%
                      arrange(desc(Total_profit))

head(Top_Country_china_df,3)

```




page 4
==============================================================================

Row
------------------------------------------------------------------------------

### Products sub categories in the US with highest average shipping cost

```{r}
US_avg_shipping_by_category <- orders %>%
                                select(Country, `Shipping Cost`, Sub_Category)%>%
                                filter(Country == "United States")%>%
                                group_by(Sub_Category) %>%
                              summarise(cost =mean(`Shipping Cost`))


US_avg_ship_bycat <-   US_avg_shipping_by_category %>%
                         data.frame()%>%
                        arrange(desc(cost))
US_avg_ship_bycat
```


```{r}
#Let's visualize
 US_avg_ship_bycat$Sub_Category <- factor(US_avg_ship_bycat$Sub_Category, levels= US_avg_ship_bycat$Sub_Category[order(US_avg_ship_bycat$cost)])

```

page 5
==============================================================================

Row
------------------------------------------------------------------------------

### Products Sub categories with highest average shipping cost in the US


```{r}

US_avg_ship_bycat <- US_avg_ship_bycat %>% as.data.frame()
```



```{r}

q2a <- ggplot(data = US_avg_ship_bycat, aes(y=reorder(Sub_Category, + cost), x= cost, fill = Sub_Category)) + geom_bar(stat = "identity") + ggtitle(label = "Plot of Products Sub categories with highest average shipping cost in the US") + geom_text(aes(label = round(cost)),hjust = -0.0008) + labs(x = "Cost", y = "Sub_category")

#q2a +theme(axis.title.x = element_text(color = "blue", size = 14, face = "bold"))

ggplotly(q2a)

```


Question 3.
a) Assess Nigeria’s profitability (i.e., total profit) for 2014. How does it compare to other
African countries?
b) What factors might be responsible for Nigeria’s poor performance? You might want to
investigate shipping costs and the average discount as potential root causes.

page 6
==============================================================================

Row
------------------------------------------------------------------------------

### Table of Profit: Nigeria Vs Rest of AFrica

```{r}
naija_vs_Africa <- orders %>%
          select(Region, Country, Profit, Discount,Year) %>%
          filter(Region == "Africa" & Year =="2014")

naija_profit <- naija_vs_Africa %>% group_by(Country) %>% summarise(Mean_profit = mean(Profit))

naija_vs_Africa <- naija_vs_Africa %>% as.data.frame()%>% arrange(desc(Profit))

naija_vs_Africa
```

page 7
=======================================================================

Row
-----------------------------------------------------------------------

###  Nigeria Vs Rest of AFrica

```{r warning=FALSE}
#Get the plot

q3 <- ggplot(data = naija_profit, aes(y= reorder(Country,+Mean_profit), x = Mean_profit, fill = Country)) + geom_bar(stat= "identity")+ ggtitle(label = "Plot of profitablity across nations") +
  geom_text(aes(label = round(Mean_profit)),hjust = -0.2)+  labs(x = "Average Profit", y = "Country")
  
  ggplotly(q3)

```

Interpretation: In Africa, on the average, Nigeria is the least profitable country.

This may be due to  a lot of factors.

Nigeria, Uganda and Zimbabwe should consider tuning down on the discount. They are the only countries with the highest discounts and accounting for loss.

<h2>possible Factors</h2>
Nigeria and Uganda and Zimbabwe should consider cutting down on discounts. They are the only with the highest discount and they are are the bottom of the profit table.

page 8
=======================================================================

Row
-----------------------------------------------------------------------

### Country Vs Discount

```{r}
possible_fac <- naija_vs_Africa %>% #arrange(desc(Discount)) %>%
  group_by(Country) %>% summarise(Mean_Discount = mean(Discount))

possible_fac <- possible_fac %>% arrange(desc(Mean_Discount))

possible_fac <- possible_fac %>% as.data.frame()

possible_fac
```

Countries and discount offered

page 9
==============================================================================

Row
------------------------------------------------------------------------------

### Least Profitable product sub category in Southeast Asia

```{r}
asia_list <- orders %>% select (Region, Country, Sub_Category, `Shipping Cost`, Profit) %>%
  filter(Region == "Southeast Asia" )

asia_least_profitable <- asia_list %>% as.data.frame() %>%
                    arrange(desc(Profit))

tail(asia_least_profitable,50)

```

The least profitable product category are Tables, Bookcases and Accessories.


```{r}
stop_asia <- asia_least_profitable %>%
                filter(Sub_Category == "Tables")%>%
               group_by(Country) %>% summarise(Avg_Profit = sum(Profit))

stop_asia <- stop_asia %>% as.data.frame()
```



page 10
=======================================================================

Row
-----------------------------------------------------------------------

### Least profitable country in Southeast Asia

```{r}
q4 <- ggplot(data = stop_asia , aes(y= reorder(Country,-Avg_Profit), x = Avg_Profit, fill = Country)) + geom_bar(stat= "identity")+ ggtitle(label = "Plot of profits per country in Asia")+
  geom_text(aes(label = round(Avg_Profit)),hjust = 1.2) + labs(x = "Average Profit", y = "Country")
  
 
ggplotly(q4)

```




-----------------------------------------------------------------------------



Question 5.
a) Which city is the least profitable (in terms of average profit) in the United States? For
this analysis, discard the cities with less than 10 Orders.
b) Why is this city’s average profit so low?




```{r}
city_usa <- orders %>%
            filter(Country == "United States") %>%
            select(City, Quantity, Profit)


#least_usa_city <- city_usa %>% filter( Quantity > 9 & City == "New York City")

least_usa_city <- city_usa %>% filter( Quantity > 9 )%>%
                    group_by(City)%>%
                    summarise(Total_Quantity = sum(Quantity),Avg_profit =   mean(Profit),  Total_pro = sum(Profit))

#View(least_usa_city)
```

page 11
=======================================================================

Row
-----------------------------------------------------------------------

### Average least profitable city in the US

```{r}
least_usa_city <- least_usa_city %>% as.data.frame() %>%                  arrange(desc(Avg_profit))

least_usa_city

```

List of profitable cities in the US, from the highest to the lowest


```{r}
#get the plot of the negative average profits.
 least_usa_city_nve <- least_usa_city %>% filter( Avg_profit < 0)

```

page 11
=======================================================================

Row
-----------------------------------------------------------------------

### least profitable city

```{r}
q5 <- ggplot(data = least_usa_city_nve, aes( y = reorder(City, -Avg_profit), x = Avg_profit, fill = City)) + geom_bar(stat ="identity") + ggtitle(label = "Plot of profits per city in USA") +  labs(x = "Average Profit", y = "Country")
  
  
ggplotly(q5)

```

Concord is the least profitable city in the United States with quantity > 9.


```{r}
concord_reason <- orders %>%
            filter(Country == "United States"  & Quantity >9) %>%
            select(City, Quantity, Profit, `Shipping Cost`, Discount)%>% group_by(City) %>% summarise(Quantity = sum(Quantity),Profit = mean(Profit),`Shipping Cost` = sum(`Shipping Cost`), Discount =sum(Discount))

concord_reason <- concord_reason %>% as.data.frame()%>% arrange(desc(`Shipping Cost`))

concord_reason_20 <- concord_reason[c(1:10),]

concord_reason_20$City  <- factor(concord_reason_20$City,                             levels = concord_reason_20$City[order(concord_reason_20$`Shipping Cost`)])
```

page 12
=======================================================================

Row
-----------------------------------------------------------------------

### Probable reason for low profit
```{r}
q5b <- ggplot(data = concord_reason_20, aes( y = City, x = `Shipping Cost`, fill = City)) + geom_bar(stat ="identity") + ggtitle(label = "Plot of shipping cost per city in the USA")+
  geom_text(aes(label = round(`Shipping Cost`)),hjust = 0.000000000951)

ggplotly(q5b)
```

Concord shipping cost is the 2nd highest in the cities of USA, this could be a probable cause for low profit margin.
-- check items, stacked barchart of shipping cost and quantity.
------------------------------------------------------------------------------


Question 6.
a) Which product subcategory has the highest average profit in Australia?

page 13
=======================================================================

Row
-----------------------------------------------------------------------

### product subcategory hwith the highest average profit in Australia 
 
```{r}
Aust_sub <- orders %>%  filter(Country == "Australia") %>%
             select(Sub_Category, Profit) %>%
              group_by(Sub_Category) %>% summarise( Avg_Profit = mean(Profit))

Aust_sub <- Aust_sub %>% as.data.frame() %>% arrange(desc(Avg_Profit))



Aust_sub$Sub_Category <- factor(Aust_sub$Sub_Category, levels = Aust_sub$Sub_Category[order(Aust_sub$Avg_Profit)])

Aust_sub
```

page 14
=======================================================================

Row
-----------------------------------------------------------------------

### product subcategory with the highest average profit in Australia 

```{r}
q56b <- ggplot(data = Aust_sub, aes( y = Sub_Category, x = Avg_Profit, fill = Sub_Category)) + geom_bar(stat ="identity")+ ggtitle(label = "Plot of product category per average profit in Australia")+
  geom_text(aes(label = round(Avg_Profit)),hjust = 0.01)+ labs(x = "Average Profit", y = "Sub_category")

ggplotly(q56b)

```

The product subcategory in Australia with the highest average profit is Appliances.



Question 7.
a)Which customer returned items and what segment do they belong
b)Who are the most valuable customers and what do they purchase?

page 15
=======================================================================

Row
-----------------------------------------------------------------------

### Customers who returned purchased items

```{r}
#extract segment, purchase, order_ID from orders table

mini_table <- orders %>% select(Segment,Order_ID,Customer_Name,Customer_ID,Product_Name, Quantity)

#A new table
customer_return <- left_join(returns,mini_table, by ='Order_ID')

customer_isonu <- customer_return %>% group_by(Customer_Name) %>% summarise(Segment,Quantity = sum(Quantity))

customer_isonu <- customer_isonu %>% distinct() %>% arrange(desc(Quantity))
customer_isonu %>% as.data.frame()

customer_isonu
```
Here are the people who returned all categories of items.

page 16
=======================================================================

Row
-----------------------------------------------------------------------

### Customers who returned purchased items Consumer category

```{r}
#break it down by Segment
#unique(customer_isonu$Segment)

customer_isonu_Consumer <- customer_isonu %>% filter(Segment == "Consumer")
customer_isonu_Consumer <- customer_isonu_Consumer %>% as.data.frame()
customer_isonu_Consumer
```
Customers who returned purchased items Consumer category

page 17
=======================================================================

Row
-----------------------------------------------------------------------

### Customers who returned purchased items Corporate category


```{r}
customer_isonu_Corporate <- customer_isonu %>% filter(Segment == "Corporate")
customer_isonu_Corporate <- customer_isonu_Corporate %>% as.data.frame()
customer_isonu_Corporate
```
Customers who returned purchased items Corporate category


page 18
=======================================================================

Row
-----------------------------------------------------------------------

### Customers who returned purchased items Home Office category


```{r}
customer_isonu_HO <- customer_isonu %>% filter(Segment == "Home Office")
customer_isonu_HO <- customer_isonu_HO %>% as.data.frame()
customer_isonu_HO

```
Customers who returned purchased items Home Office category


```{r}
#Lets visualize
 new_view <- customer_isonu %>% group_by(Segment) %>% summarise(Freq = n())
```

page 19
=======================================================================

Row
-----------------------------------------------------------------------

### returned items per category

```{r}

ggplot( data = new_view, aes(x= Segment, y = Freq, fill= Segment))+ geom_bar(stat="identity") + ggtitle(label = "Plot of returned goods per segment")+
  geom_text(aes(label = round(Freq)), h=.65)+labs(x = "Segment", y = "Count of returned Items")

#customer_isonu_Consumer
#customer_isonu_Corporate
#customer_isonu_HO
```




```{r}
MVPk <- mini_table %>% group_by(Customer_ID) %>% summarise(Product_Name,Customer_Name, Quantity = sum(Quantity))

MVPL <- MVPk %>% as.data.frame( ) %>% arrange(desc(Quantity))

```
Top Customers and the products they bought!

page 20
=======================================================================

Row
-----------------------------------------------------------------------

### Most Valuable Customer


```{r}
#lets visualize customers per quantity ordered
visualize_MVP <- MVPL %>% select(-Product_Name) %>%distinct(Customer_Name, Quantity )

```


```{r echo=FALSE, paged.print=TRUE}

#subset the data
rmvp <- visualize_MVP[c(1:30),]
rmvp <- rmvp %>% as.data.frame( ) %>% arrange(desc(Quantity))


q7 <- ggplot(data = rmvp, aes(y = reorder(Customer_Name, +Quantity ), x= Quantity, fill = Customer_Name)) + geom_bar(stat= "identity")+ ggtitle(label = "Plot of returned goods per segment")+
  geom_text(aes(label = round(Quantity)),hjust = 0.012) + labs(x = "Quantity", y = "Customer Name")

ggplotly(q7)


```

page 21
=======================================================================

Row
-----------------------------------------------------------------------

### Most Valuable Customer

```{r}
MVP <- mini_table %>% group_by(Customer_ID) %>% summarise(Product_Name,Customer_Name, Quantity = sum(Quantity))

MVP <- MVP %>% as.data.frame( ) %>% arrange(desc(Quantity))
head(MVP,100)

```
Customers and what they purchased. 

